---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE  # Set to TRUE when you have real data to run examples
)
```

# illibindex

<!-- badges: start -->
[![R-CMD-check](https://github.com/FraWagner/illibindex/workflows/R-CMD-check/badge.svg)](https://github.com/FraWagner/illibindex/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

## Overview

**illibindex** measures liberal and illiberal rhetoric in political discourse using word embeddings and cosine similarity. The package helps researchers quantify how political actors position themselves on policy issues by comparing their language to predefined liberal and illiberal vocabularies.

### Key Features

- 📚 **Pre-built dictionaries** for immigration, gender, and other policy domains
- 🧠 **Word embedding training** using GloVe with bootstrap resampling
- 📊 **Cosine similarity analysis** to measure ideological positioning
- 🎯 **Flexible workflow** from text to validated measurements

## Installation

Install the development version from GitHub:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("FraWagner/illibindex")
```

## Quick Example

Here's a simple workflow showing how to use the package:

```{r quick-example}
library(illibindex)

# 1. Load built-in dictionaries
data(dictionaries)

# 2. Process the immigration dictionary
immigration_dict <- process_dictionary(dictionaries, "immigration")

# 3. View available terms
immigration_dict$terms
#> [1] "immigra" "border"  "asylum"  "migrat"  "migrant"

# 4. Examine liberal terms for "immigra"
head(immigration_dict$liberal$immigra)
#> [1] "welcome"     "diversity"   "integrate"   "opportunity" 
#> [5] "refuge"      "compassion"

# 5. Examine illiberal terms for "immigra"
head(immigration_dict$illiberal$immigra)
#> [1] "illegal"   "invasion"  "threat"    "criminal"  
#> [5] "deport"    "close"
```

## Full Workflow

### Step 1: Prepare Your Text Data

Your corpus should be organized with metadata identifying the political actor and time period:

```{r data-prep}
# Example structure (your actual data)
corpus_df <- data.frame(
  text = c("We must welcome refugees...", "Close the borders..."),
  actor = c("Party_A", "Party_B"),
  time = c("2020", "2020"),
  country = c("Italy", "Italy")
)
```

### Step 2: Train Word Embeddings

Train domain-specific word embeddings for each actor and time period:

```{r train-embeddings}
# Train embeddings with bootstrap resampling
train_word_embeddings(
  country = "Italy",
  model = "model1",
  output_dir = "output/embeddings",
  n_bootstrap = 100,        # Number of bootstrap samples
  window_size = 6,          # Context window
  dim = 300                 # Embedding dimensions
)

# This creates files like:
# output/embeddings/Italy/model1/wordvecs_PartyA_2020.rds
# output/embeddings/Italy/model1/wordvecs_PartyB_2020.rds
```

### Step 3: Calculate Cosine Similarity

Measure how close each actor's language is to liberal vs. illiberal terms:

```{r calc-similarity}
# Calculate similarity scores
results <- calc_cosSim(
  country = "Italy",
  model = "model1",
  which = "immigration",
  embeddings_path = "output/embeddings/Italy/model1",
  save = TRUE,
  output_dir = "output/results"
)

# View results structure
str(results)
```

### Step 4: Analyze Results

The output includes mean similarity scores with confidence intervals:

```{r analyze-results}
# Extract results
immigration_results <- results$immigration

# Liberal-Illiberal difference scores (higher = more liberal positioning)
head(immigration_results)
#>   mean     sd  lowerci upperci dimension   Policy  Party Time    word
#> 1 0.12  0.045    0.10    0.14  LibIllib immigration PartyA 2020 immigra
#> 2 -0.08 0.052   -0.10   -0.06  LibIllib immigration PartyB 2020 immigra

# Visualize positioning over time
library(ggplot2)

ggplot(immigration_results[immigration_results$dimension == "LibIllib", ], 
       aes(x = word, y = mean, color = Party)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lowerci, ymax = upperci), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Liberal-Illiberal Positioning on Immigration",
    x = "Policy Term",
    y = "Liberal-Illiberal Score",
    caption = "Positive = more liberal, Negative = more illiberal"
  ) +
  theme_minimal() +
  coord_flip()
```

## Available Dictionaries

The package includes dictionaries for multiple policy domains:

```{r dictionaries}
data(dictionaries)
names(dictionaries)
#> [1] "immigration" "gender" ...

# Each dictionary contains liberal and illiberal term lists
# organized by policy-relevant concepts
```

## Understanding the Scores

**Liberal-Illiberal (LibIllib) Score:**
- Calculated as: `liberal_similarity - illiberal_similarity`
- **Positive values**: Actor's language closer to liberal vocabulary
- **Negative values**: Actor's language closer to illiberal vocabulary
- **Near zero**: Neutral or ambiguous positioning

**Individual Dimensions:**
- `liberal`: Mean cosine similarity to liberal terms
- `illiberal`: Mean cosine similarity to illiberal terms
- `LibIllib`: The difference score (main measure of interest)

## Advanced Usage

### Custom Dictionaries

You can create your own dictionaries following the package format:

```{r custom-dict}
# Create a custom dictionary data frame
custom_dict <- data.frame(
  country = "all",
  term = NA,
  climate_liberal = c("renewable", "sustainable", "green", NA),
  climate_illiberal = c("hoax", "expensive", "unnecessary", NA),
  stringsAsFactors = FALSE
)

# Add to dictionaries list
dictionaries$climate <- custom_dict

# Process it
climate_dict <- process_dictionary(dictionaries, "climate")
```

### Filtering by Country

If your dictionary includes country-specific terms:

```{r country-filter}
# The process_dictionary function filters to country == "all" by default
# Modify the function call to include specific countries as needed
```

## Citation

If you use this package in your research, please cite:

```
Wagner, F. (2024). illibindex: Measuring Liberal and Illiberal Rhetoric 
in Political Discourse. R package version 0.1.0. 
https://github.com/FraWagner/illibindex
```

## Related Work

This package implements methods for measuring ideological positioning using word embeddings. For theoretical background and validation, see:

- ["Measuring Illiberalism: Mapping Illiberalism in Seven Countries, 2000-2022."](https://preprints.apsanet.org/engage/apsa/article-details/67840ffc6dde43c9083af8cb) 
- ["Opposition to Government and Back: How Illiberal Parties Shape Immigration Discourse and Party Competition"](https://www.cogitatiopress.com/politicsandgovernance/article/view/9609)


---