---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# illibindex

<!-- badges: start -->
<!-- badges: end -->

The goal of illibindex is to provide a simple way to implement the illiberalism index as used in the paper ["Measuring Illiberalism: Mapping Illiberalism in Seven Countries, 2000-2022."](https://preprints.apsanet.org/engage/apsa/article-details/67840ffc6dde43c9083af8cb) 

## Installation

You can install the development version of illibindex from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("FraWagner/illibindex")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(illibindex)
## basic example code
```

## Citation

Please cite the package using the [pre print](https://preprints.apsanet.org/engage/apsa/article-details/67840ffc6dde43c9083af8cb) DOI:  <10.33774/apsa-2025-xdd54-v2> 
